Engineering the Servo Web Browser Engine using Rust

### 2015-10-24 icse16-servo/abstract.tex
Abstract & intro fixes

In this paper, we show how a language with an advanced type system can address many of
the most common security issues and software engineering challenges in other browser
engines, while still producing code that has the same performance and memory profile
required by these applications.
This language is also quite accessible to new open source contributors and employees,
even those without a background in \Cplusplus{} or systems programming.
We also outline several pitfalls encountered along the way and describe some potential
areas for future improvement.
\end{abstract}

>>>>

In this paper, we show how a language with an advanced type system can address many of
the most common security issues and software engineering challenges in other browser
engines, while still producing code that has the same performance and memory profile.
This language is also quite accessible to new open source contributors and employees,
even those without a background in \Cplusplus{} or systems programming.
We also outline several pitfalls encountered along the way and describe some potential
areas for future improvement.
\end{abstract}

### 2015-10-24 icse16-servo/intro.tex
Abstract & intro fixes

The heart of a modern web browser is its engine, the code responsible
for loading, processing, evaluating, and rendering web content.
There are three major browser engine families:
\begin{enumerate}
\item Trident/Spartan, the engine in Internet Explorer~\cite{IE}
\item Webkit\cite{WEBKIT}/Blink, the engine in Safari~\cite{SAFARI}, Chrome~\cite{CHROME}, and Opera~\cite{OPERA}
\item Gecko, the engine in Firefox~\cite{FIREFOX}
\end{enumerate}
All of these engines have at their core many millions of lines of \Cplusplus{} code.
The use of \Cplusplus{} has enabled all of these browsers to achieve excellent sequential
performance on a single web page, particularly on desktop computers.
But, they all face several challenges:
\begin{itemize}
  \item On mobile devices with lower processor speed but many more processors, these browsers
    do not provide the same level of interactivity~\cite{parallelizing-web-pages,ZOOMM}.
  \item In Gecko, we determined that roughly 50\% of the security critical bugs are memory use after
    free, array out of range access, or related to integer overflow.
  \item As the web has become more interactive, the mostly-sequential architecture of these engines
    has made it challenging to incorporate new features without sacrificing interactivity.
  \item With the growth in the popularity of other languages at the expense of \Cplusplus{}, both
    the number of volunteer contributors to the core \Cplusplus{} parts of these browser engine
    open source codebases has not grown apace with the increase in the size
    of the codebase.
\end{itemize}

@@

To address memory-related safety issues, we are using a new systems programming language,
Rust~\cite{RUST}.
In Rust, errors such as off-by-one array access or memory buffer use after free are
prevented statically at compile time.

@@

Servo is currently over 800k lines of Rust code and implements enough of the web platform to render and
process many pages, though it is still a far cry from the over 7 million lines of code in
the Mozilla Firefox browser.
However, we believe that we have implemented enough of the web platform to provide an
early report on the successes, failures, and open problems remaining in Servo, from the
point of view of programming languages and runtime research.
In this experience report, we discuss the design and architecture of a modern web 
browser engine, show how using the Rust programming language has helped us to  address the
engineering challenges we have encountered when building the browser engine, and also touch on
open problems and areas of future investigation.

>>>>

The heart of a modern web browser is its engine, the code responsible
for loading, processing, evaluating, and rendering web content.
There are three major browser engine families:
\begin{enumerate}
\item Trident/Spartan, the engine in Internet Explorer~\cite{IE}
\item Webkit\cite{WEBKIT}/Blink, the engine in Safari~\cite{SAFARI}, Chrome~\cite{CHROME}, and Opera~\cite{OPERA}
\item Gecko, the engine in Firefox~\cite{FIREFOX}
\end{enumerate}
All of these engines have at their core many millions of lines of \Cplusplus{} code.
The use of \Cplusplus{} has enabled all of these browsers to achieve excellent sequential
performance on a single web page, particularly on desktop computers.
But, they all face several challenges:
\begin{itemize}
  \item On mobile devices with lower processor speed but many more processors, these browsers
    do not provide the same level of interactivity~\cite{parallelizing-web-pages,ZOOMM}.
  \item In Gecko, roughly 50\% of the security critical bugs are memory use after free, array
    out of range access, or related to integer overflow, all mistakes commonly made by even
    experienced C++ programmers with access to the best static analysis tools available.
  \item As the web has become more interactive, the mostly-sequential architecture of these engines
    has made it challenging to incorporate new features without sacrificing interactivity.
  \item With the growth in the popularity of other languages at the expense of \Cplusplus{}, 
    the number of volunteer contributors to the core \Cplusplus{} parts of these browser engine
    open source codebases has not grown apace with the increase in the size
    of the codebase.
\end{itemize}

@@

To address memory-related safety issues, we are using a new systems programming language,
Rust~\cite{RUST}.
In Rust, errors such as off-by-one array access or memory buffer use after free are
prevented by the language and its builtin libraries.

@@

Servo is currently over 800k lines of Rust code and implements enough of the web platform to render and
process many pages, though it is still a far cry from the over 7 million lines of code in
the Mozilla Firefox browser.
However, we believe that we have implemented enough of the web platform to provide an
early report on the successes, failures, and open problems remaining in Servo, from the
point of view of experimenting with the new programming language, Rust.
In this experience report, we discuss the design and architecture of a modern web 
browser engine, show how using the Rust programming language has helped us to address the
engineering challenges we have encountered when building the browser engine, and also touch on
open problems and areas of future investigation.
